\hspace{1cm} Do syntactic and semantic similarity lead to interference effects? \newline Evidence from self-paced reading and event-related potentials using German

### 2024-01-05 paper/main.tex
updated uniform prior

    \begin{subfigure}[t]{\textwidth}
        \centering
         \caption{On the microvolt scale, i.e., for ERPs.}\label{fig:model_predictions_ERP}
         \includegraphics[width=\textwidth]{images/ERP_priorpred_vs_postpred_vs_data.png}
     \end{subfigure}
\end{figure}

The \citet{Lewis2005} model is generally used to predict reading times. The scaling factor $F$ links the activation of items in memory to retrieval times \citep{Lewis2005,anderson_etal_2004_ACT_R}. But this scaling factor can be adjusted to link activation to brain responses to use the model to generate predictions on the microvolt scale for ERP data. The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data. To generate prior predictions on the microvolt scale, we sampled the free scaling parameter $F$ from a truncated normal prior distribution ($Normal_+(0.01, 0.005)$) with the lower and upper bounds at 0.003 and 0.04, respectively. As in the reading time modeling, to avoid overfitting we fitted $F$ only to the semantic interference effect in order to generate posterior predictive distributions of the syntactic and semantic interference effects. For the ERP modeling, 3,146 out of 20,000 sampled values of $F$ lead to simulated effects within the range of the observed effect and were thus used for posterior predictions. The prior and posterior predictions for the effects of interest on the microvolt scale in comparison to our ERP data are shown in Figure \ref{fig:model_predictions_ERP}. The prior and posterior predictions of the model overlap with the effects of interest in the data. The general effect pattern (no syntactic interference effect, but a semantic interference effect) is similar in both the prior and posterior predictive distributions. Therefore, the assumption that the relevant syntactic cue is \{$\pm$ subject-in-same-clause\} cue, rather than  the \{$\pm$ grammatical subject\} cue originally assumed in previous modeling work, seems reasonable for the present ERP study as well.

>>>>

    \begin{subfigure}[t]{\textwidth}
        \centering
         \caption{On the microvolt scale, i.e., for negative ERPs.}\label{fig:model_predictions_ERP}
         \includegraphics[width=\textwidth]{images/ERP_priorpred_vs_postpred_vs_data.png}
     \end{subfigure}
\end{figure}

The \citet{Lewis2005} model is generally used to predict reading times. The scaling factor $F$ links the activation of items in memory to retrieval times \citep{Lewis2005,anderson_etal_2004_ACT_R}. But this scaling factor can be adjusted to link activation to brain responses to use the model to generate predictions on the microvolt scale for ERP data. The modeling reported here is, to the best of our knowledge, the first time that the \citet{Lewis2005} model has been fit quantitatively to ERP data. We sampled the free scaling parameter $F$ from a uniform prior distribution ($Uniform(0.005, 0.04)$). This prior reflects the a priori belief that effects will range from 0.5 to 4 $\mu V$ with equal probability for all effect sizes in this range. We are modeling N400 amplitude differences, so larger predicted effect sizes correspond to a more negative N400. To reflect this, the predicted effects were multiplied with $-1$. As in the reading time modeling, to avoid overfitting we fitted $F$ only to the semantic interference effect in order to generate posterior predictive distributions of the syntactic and semantic interference effects. For the ERP modeling, 473 out of 20,000 sampled values of $F$ lead to simulated effects within the range of the observed effect and were thus used for posterior predictions. The prior and posterior predictions for the effects of interest on the microvolt scale in comparison to our ERP data are shown in Figure \ref{fig:model_predictions_ERP}. The prior and posterior predictions of the model overlap with the effects of interest in the data. The general effect pattern (no syntactic interference effect, but a semantic interference effect) is similar in both the prior and posterior predictive distributions. Therefore, the assumption that the relevant syntactic cue is \{$\pm$ subject-in-same-clause\} cue, rather than the \{$\pm$ grammatical subject\} cue originally assumed in previous modeling work, seems reasonable for the present ERP study as well.
